---
title: "Slurm Simulator: Micro Cluster Tutorial"
author: nikolays@buffalo.edu
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    toc: yes
    toc_float: yes
    toc_depth: 4
    mathjax: null
    css: doc.css
    df_print: paged

editor_options: 
  markdown: 
    wrap: 80
---

# Slurm Simulator: Micro Cluster Tutorial

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load R libraries (warnings on masking are ok)
library(tidyverse)
library(plotly)
library(cowplot)
library(devtools)
library(reticulate)
library(knitr)
use_condaenv(condaenv="base")

# rebuild RSlurmSimTools
load_all("/opt/slurm_sim_tools/src/RSlurmSimTools")
# library(RSlurmSimTools)
```

```{bash}
# Check that MySQL Server is up
ps -Af | grep mariadbd 
```

## Overview

This tutorial will teach you how to run a Slurm Simulator using an example of a
small 10-node Micro-Cluster.

The tutorial will walk you through the Slurm simulation configuration.

You will run two simulations with identical workloads but different timing
between the start-up and first job to investigate scheduling stochasticity.

Finally, you'll run ten independent simulations with the same workload and
calculate statistics other than runs.

The files needed for this tutorial are located at
`slurm_sim_tools/tutorials/micro_cluster`.

## Micro-Cluster Description

In this tutorial small model cluster, named Micro-Cluster, will be simulated.
Micro-Cluster is created to test various aspects of the slurm simulator usage.
It consists of 10 compute nodes (see Table below). There are two types of
general compute nodes with different CPU types to test features request
(constrains), one GPU node to test GRes and one big memory node to test proper
allocation of jobs with large memory requests.

| Node Type       | Node Name | Number of Cores | CPU Type | Memory Size | Number of GPUs |
|-----------------|-----------|-----------------|----------|-------------|----------------|
| General Compute | n[1-4]    | 12              | CPU-N    | 48GiB       | -              |
| General Compute | m[1-4]    | 12              | CPU-M    | 48GiB       | -              |
| GPU node        | g1        | 12              | CPU-G    | 48GiB       | 2              |
| Big memory      | b1        | 12              | CPU-G    | 512GiB      | -              |

: Micro-Cluster compute nodes description

The nodes are connected as follows:

```         
Top Switch
 ├─IBN Switch
 │  ├─Node n1
 │  ├─Node n2
 │  ├─Node n3
 │  └─Node n4
 ├─IBM Switch
 │  ├─Node m1
 │  ├─Node m2
 │  ├─Node m3
 │  └─Node m4
 └─IBG Switch
    ├─Node g1
   `└─Node b1


```

## Slurm Simulation Configuration

The UB Slurm Simulator is made from Slurm source code and so it shares many
configurations with actual Slurm installation. In addition, you also need to
specify that simulated users will submit to the simulated cluster and some
simulation parameters.

In order to run Slurm simulation you need:

1.  Adopt the Slurm configuration (`slurm.conf` and friends) for simulation
    -   Some options are not supported and files, typically used in slurm need
        to be renamed and point to user space.
2.  Set-up users and accounts: create `sacctmgr.script` and `user.sim`.
3.  Specify simulation parameters
    -   `sim.conf` some of parameters needed to be reset at run-time
4.  Create events file with workload specification (a.k.a. job-traces or
    events-list).

```{bash}
ls etc
```

### Adopting the Slurm Configuration for Simulation

The simulator is built from Slurm code and uses regular Slurm configuration.

In the `etc` directory, you can find several familiar Slurm configuration files:
`slurm.conf`, `topology.conf`, `gres.conf` and `slurmdbd.conf`.

When you start from the actual cluster configuration, some modifications are
needed to point to different file system locations and disable some features
which the simulator cannot work with.

The files in this tutorial are already modified to be executed in simulator
mode. In the following sub-section, we will list those changes.

#### Changes in `slurm.conf`

```ini
# (optional) Different cluster name might be useful
ClusterName=micro

# Authentication is turned off
AuthType=auth/none

# This user will be used to run slurm simulator
SlurmUser=slurm
# (not needed) UB Slurm simulator does not use slurmd
SlurmdUser=root
# (not needed) This parameter is for Virtual Clusters (not used in Simulator)
SlurmdParameters=config_overrides

# Change Slurm control daemon address to localhost
ControlMachine=localhost
ControlAddr=localhost
# Change Slurm Database daemon address to localhost
AccountingStorageHost=localhost

# Change file location
JobCredentialPrivateKey=/home/slurm/work/micro_cluster/etc/slurm.key
JobCredentialPublicCertificate=/home/slurm/work/micro_cluster/etc/slurm.cert
SlurmSchedLogFile=/home/slurm/work/micro_cluster/log/sched.log
SlurmctldLogFile=/home/slurm/work/micro_cluster/log/slurmctld.log
SlurmdLogFile=/home/slurm/work/micro_cluster/log/slurmd.log
SlurmdSpoolDir=/home/slurm/work/micro_cluster/var/spool
StateSaveLocation=/home/slurm/work/micro_cluster/var/state
JobCompLoc=/home/slurm/work/micro_cluster/log/jobcomp.log

# Have to set it as simulator uses front-end mode
FrontEndName=localhost
```

#### Changes in `slurmdbd.conf`

```ini
# Change Slurm Database daemon address to localhost
DbdHost=localhost
# Authentication is turned off
AuthType=auth/none

# Change file location
PidFile=/home/slurm/work/micro_cluster/var/slurmdbd.pid
LogFile=/home/slurm/work/micro_cluster/log/slurmdbd.log

# Change mysql server address to localhost and accounts
StorageHost=localhost
StorageUser=slurm
StoragePass=slurm
# This user will be used to run slurm dbd
SlurmUser=jovyan
# (optional) database name in mysql to use
StorageLoc=slurmdb_micro
```

#### Changes to `gres.conf`

```ini
# There is no actual hardware everything should be specified manually
AutoDetect=off
```

#### Changes to `topology.conf`

It should be fine as is.

### Set up Users, Accounts, QoS Specs

Usually system administrator will create user and accounts on the HPC resource
on per need base. In simulator, we need to create a file with simulated system
usernames and a script for Slurm `sacctmgr` utility to initiate cluster QoS-es,
users and accounts.

#### System User names and Groups - `users.sim`

`users.sim` fount in `etc` folder specify simulated system users and group names
and ids. The file format is one line per user with colon (`:`) separated four
fields:

```         
<user named>:<user id>:<group name>:<group id>
```

The used `users.sim` file is as follows:

```         
admin:1000:admin:1000
user1:1001:group1:1001
user2:1002:group2:1002
user3:1003:group3:1003
user4:1004:group2:1002
user5:1005:group1:1001
```

#### Adding Users, Accounts and QoS - `sacctmgr.script`

`sacctmgr.script` is a script which will be executed by `sacctmgr` before the
simulation. It adds slurm users, accounts, QoS and all other things sys-admins
will do with sacctmgr.

Below is a listing of used `sacctmgr.script`:

``` shell
# add/modify QOS
modify QOS set normal Priority=0
add QOS Name=supporters Priority=100
# add cluster
add cluster Name=micro Fairshare=1 QOS=normal,supporters
# add accounts
add account name=account0 Fairshare=100
add account name=account1 Fairshare=100
add account name=account2 Fairshare=100
# add admin
add user name=admin DefaultAccount=account0 MaxSubmitJobs=1000 AdminLevel=Administrator
# add users
add user name=user1 DefaultAccount=account1 MaxSubmitJobs=1000
add user name=user2 DefaultAccount=account1 MaxSubmitJobs=1000
add user name=user3 DefaultAccount=account1 MaxSubmitJobs=1000
add user name=user4 DefaultAccount=account2 MaxSubmitJobs=1000
add user name=user5 DefaultAccount=account2 MaxSubmitJobs=1000
# add users to qos level
modify user set qoslevel="normal,supporters"

# check results
list associations format=Account,Cluster,User,Fairshare tree withd
```

The script starts with modification of `normal` QoS and adding new QoS,
`priority`, with higher priority factor.

Then, we add Slurm users and accounts. Slurm user names should match the system
user names we specified at `users.sim`.

Here we have 5 users (user1, user2, user3, user4, user5) grouped into 2 accounts
(account1, account2). There is also an admin user with an associated account. We
allowed all users to use both QoS-es.

Finally, we printed created association (relationship between users and
accounts)

## Simulation Parameters

In the etc directory, you can find several configuration files:

``` ini
# Simulation Configuration

## Unix time stamp for the simulated start time
TimeStart = 1641013200.0
## Unix time stamp for the simulated stop time, 0 - run till all jobs are finished
TimeStop = 0
## Time between simulator start and first job execution (can be overwritten by `slurmsim run_sim`)
SecondsBeforeFirstJob = 126
## CPU clock scaling
ClockScaling = 1.0
## Events File (can be overwritten by `slurmsim run_sim`)
EventsFile = /home/slurm/work/micro_cluster/job_traces/jobs500_shrinked.events
## Seconds after all events are done, -1 keep spinning time_after_all_events_done.
TimeAfterAllEventsDone = 10

## Some time delays to mimic start-up times of actual Slurm controller
FirstJobDelay = -0.32
CompJobDelay = 0.000
TimeLimitDelay = 0.000
```

Some of these parameters can be overwritten from the command line

## Workload Specification (events file, a.k.a. job traces)

In `job_traces` you can find many files finished with `.events` extension, these
are events list file which used to specify users activities on simulated
cluster.

Here is the listing of `small.events`:

``` bash
-dt 0 -e submit_batch_job | -J jobid_1001 -sim-walltime 0 --uid=user5 -t 00:01:00 -n 12 --ntasks-per-node=12 -A account2 -p normal -q normal pseudo.job
-dt 1 -e submit_batch_job | -J jobid_1002 -sim-walltime -1 --uid=user1 -t 00:01:00 -n 1 --ntasks-per-node=1 -A account1 -p normal -q normal --constraint=CPU-N pseudo.job
-dt 2 -e submit_batch_job | -J jobid_1003 -sim-walltime 5 --uid=user4 -t 00:01:00 -n 1 --ntasks-per-node=1 -A account2 -p normal -q normal --mem=500000 pseudo.job
-dt 16 -e submit_batch_job | -J jobid_1004 -sim-walltime 21 --uid=user3 -t 00:01:00 -n 24 --ntasks-per-node=12 -A account1 -p normal -q normal pseudo.job
-dt 19 -e submit_batch_job | -J jobid_1005 -sim-walltime 2 --uid=user5 -t 00:01:00 -n 12 --ntasks-per-node=12 -A account2 -p normal -q normal --mem=500000 pseudo.job
-dt 19 -e submit_batch_job | -J jobid_1006 -sim-walltime 9 --uid=user3 -t 00:01:00 -n 48 --ntasks-per-node=12 -A account1 -p normal -q normal pseudo.job
-dt 19 -e submit_batch_job | -J jobid_1007 -sim-walltime -1 --uid=user4 -t 00:01:00 -n 24 --ntasks-per-node=12 -A account2 -p normal -q normal --constraint=CPU-M pseudo.job
-dt 22 -e submit_batch_job | -J jobid_1008 -sim-walltime 0 --uid=user4 -t 00:01:00 -n 12 --ntasks-per-node=12 -A account2 -p normal -q normal --constraint=CPU-M pseudo.job
-dt 26 -e submit_batch_job | -J jobid_1009 -sim-walltime 2 --uid=user1 -t 00:01:00 -n 96 --ntasks-per-node=12 -A account1 -p normal -q normal pseudo.job
-dt 26 -e submit_batch_job | -J jobid_1010 -sim-walltime 0 --uid=user5 -t 00:01:00 -n 12 --ntasks-per-node=12 -A account2 -p normal -q normal --constraint=CPU-N pseudo.job
-dt 29 -e submit_batch_job | -J jobid_1011 -sim-walltime 0 --uid=user4 -t 00:01:00 -n 1 --ntasks-per-node=1 -A account2 -p normal -q normal --gres=gpu:1 pseudo.job
-dt 32 -e submit_batch_job | -J jobid_1012 -sim-walltime -1 --uid=user5 -t 00:01:00 -n 1 --ntasks-per-node=1 -A account2 -p normal -q normal pseudo.job
-dt 36 -e submit_batch_job | -J jobid_1013 -sim-walltime 0 --uid=user2 -t 00:01:00 -n 1 --ntasks-per-node=1 -A account1 -p normal -q normal --mem=500000 pseudo.job
-dt 36 -e submit_batch_job | -J jobid_1014 -sim-walltime 7 --uid=user5 -t 00:01:00 -n 24 --ntasks-per-node=12 -A account2 -p normal -q normal --constraint=CPU-N pseudo.job
-dt 39 -e submit_batch_job | -J jobid_1015 -sim-walltime 18 --uid=user2 -t 00:01:00 -n 6 --ntasks-per-node=6 -A account1 -p normal -q normal pseudo.job
-dt 40 -e submit_batch_job | -J jobid_1016 -sim-walltime 25 --uid=user1 -t 00:01:00 -n 2 --ntasks-per-node=2 -A account1 -p normal -q normal --gres=gpu:2 pseudo.job
-dt 42 -e submit_batch_job | -J jobid_1017 -sim-walltime 1 --uid=user1 -t 00:01:00 -n 48 --ntasks-per-node=12 -A account1 -p normal -q normal --constraint=CPU-N pseudo.job
-dt 42 -e submit_batch_job | -J jobid_1018 -sim-walltime 0 --uid=user3 -t 00:01:00 -n 12 --ntasks-per-node=12 -A account1 -p normal -q normal pseudo.job
-dt 43 -e submit_batch_job | -J jobid_1019 -sim-walltime 34 --uid=user4 -t 00:01:00 -n 12 --ntasks-per-node=12 -A account2 -p normal -q normal --gres=gpu:2 pseudo.job
-dt 43 -e submit_batch_job | -J jobid_1020 -sim-walltime 14 --uid=user1 -t 00:01:00 -n 1 --ntasks-per-node=1 -A account1 -p normal -q normal --constraint=CPU-N pseudo.job
```

Each line corresponds to a single event, the arguments till the first pipe
symbol (`|`) correspond to event time and event type, and arguments after the
pipe symbol (`|`) correspond to event parameters. For the user's job submission
format, the arguments are just like you submit to the `sbatch` command (it is
processed by the same function), there are some additional arguments and some
normal `sbatch` arguments which have to follow certain format. Bellow is the
format:

```         
-dt <time to submit since the start of slurm controller> -e submit_batch_job | -J jobid_<jobid> -sim-walltime <walltime in seconds> --uid=<user> -t <requested time> -A <account> <Other Normal Slurm Arguments> pseudo.job
```

Here is a list of some often used `sbatch` arguments and simulator-augmented

-   `-J jobid_<jobid>`: `-J` is a normal Slurm argument to specify jobname, we
    use it for results processing automation, so give all jobs names like
    `jobid_<jobid>`, where `<jobid>` is numeric job-id. Some-times your
    simulation can be misconfigured and automatically assigned numeric `job-id`
    can be misalligned.
-   `-sim-walltime <walltime in seconds>`: specify walltime for job to run.
    Value `0` means job have to be killed by Slurm due to running outside of
    time-limit.
-   `--uid=<user>`: it is a normal Slurm argument but we is it to specify which
    user submit job
-   `-t <requested time>`: it is a normal Slurm argument for time request
-   `-A <account>`: it is a normal Slurm argument for specifying which account
    to use
-   `<Other normal Slurm arguments>`:
    -   `-n <nomber of noded>`
    -   `--ntasks-per-node=<ntasks-per-node>`
    -   `--gres=<gres request>`
    -   `--constraint=<features request>`
    -   `--mem=<memory request>`
    -   Try other `sbatch` arguments, they might work as well
-   `pseudo.job`: should finished with script name, this also used in VC

```{bash}
pwd
```

## Run First Simulations

`slurmsim` is a utility for various Slurm simulation tasks. To run the
simulation, we will use the `run_sim` subcommand. Execute following:

```bash
export CLUS_DIR=$(pwd)
export MACHINE_NAME="slurmsimcont"
export RUN_NAME="test1"
export dtstart=59
export replica=1

slurmsim -v run_sim  -d \
            -e ${CLUS_DIR}/etc \
            -a ${CLUS_DIR}/etc/sacctmgr.script \
            -w ${CLUS_DIR}/workload/small.events \
            -r ${CLUS_DIR}/results/${MACHINE_NAME}/${RUN_NAME}/dtstart_${dtstart}_${replica} \
            -dtstart $dtstart
```

```         
    Logger initialization
    [INFO] Note: NumExpr detected 16 cores but "NUMEXPR_MAX_THREADS" not set, so enforcing safe limit of 8.
    [INFO] NumExpr defaulting to 8 threads.
    [INFO] Read from /home/slurm/work/micro_cluster/job_traces/small.events 20 even ts
    [INFO] slurm.conf: /home/slurm/work/micro_cluster/etc/slurm.conf
    [INFO] slurmdbd: /opt/slurm_sim/sbin/slurmdbd
    [INFO] slurmd: /opt/slurm_sim/sbin/slurmd
    [INFO] slurmctld: /opt/slurm_sim/sbin/slurmctld
    [INFO] dropping db from previous runs
    DROP DATABASE IF EXISTS slurmdb_micro
    [INFO] directory (/home/slurm/work/micro_cluster/log) does not exist, creating it 
    [INFO] deleting previous SlurmdbdPidFile file: /home/slurm/work/micro_cluster/var/slurmdbd.pid
    [INFO] deleting previous StateSaveLocation files from /home/slurm/work/micro_cluster/var/state
    [DEBUG] Set stdout/stderr for slurmctld to /home/slurm/work/micro_cluster/log/slurmctld_stdout.log
    [DEBUG] Set stdout/stderr for slurmdbd to /home/slurm/work/micro_cluster/log/slurmdbd_stdout.log
    [INFO] Launching slurmdbd
    [INFO] Running sacctmgr script from /home/slurm/work/micro_cluster/etc/sacctmgr.script
    sacctmgr: sacctmgr:  Modified qos...
      normal
    sacctmgr:  Adding QOS(s)
      supporters
     Settings
      Description    = supporters
      Priority                 = 100
    sacctmgr: sacctmgr:  Adding Cluster(s)
      Name           = micro
     Setting
      Default Limits:
      Fairshare     = 1
      QOS           = normal,supporters
    sacctmgr: sacctmgr:  Adding Account(s)
      account0
     Settings
      Description     = Account Name
      Organization    = Parent/Account Name
     Associations
      A = account0   C = micro     
     Settings
      Fairshare     = 100
      Parent        = root
    sacctmgr:  Adding Account(s)
      account1
     Settings
      Description     = Account Name
      Organization    = Parent/Account Name
     Associations
      A = account1   C = micro     
     Settings
      Fairshare     = 100
      Parent        = root
    sacctmgr:  Adding Account(s)
      account2
     Settings
      Description     = Account Name
      Organization    = Parent/Account Name
     Associations
      A = account2   C = micro     
     Settings
      Fairshare     = 100
      Parent        = root
    sacctmgr: sacctmgr:  Adding User(s)
      admin
     Settings =
      Default Account = account0
      Admin Level     = Administrator
     Associations =
      U = admin     A = account0   C = micro     
     Non Default Settings
      MaxSubmitJobs = 1000
    sacctmgr: sacctmgr:  Adding User(s)
      user1
     Settings =
      Default Account = account1
     Associations =
      U = user1     A = account1   C = micro     
     Non Default Settings
      MaxSubmitJobs = 1000
    sacctmgr:  Adding User(s)
      user2
     Settings =
      Default Account = account1
     Associations =
      U = user2     A = account1   C = micro     
     Non Default Settings
      MaxSubmitJobs = 1000
    sacctmgr:  Adding User(s)
      user3
     Settings =
      Default Account = account1
     Associations =
      U = user3     A = account1   C = micro     
     Non Default Settings
      MaxSubmitJobs = 1000
    sacctmgr:  Adding User(s)
      user4
     Settings =
      Default Account = account2
     Associations =
      U = user4     A = account2   C = micro     
     Non Default Settings
      MaxSubmitJobs = 1000
    sacctmgr:  Adding User(s)
      user5
     Settings =
      Default Account = account2
     Associations =
      U = user5     A = account2   C = micro     
     Non Default Settings
      MaxSubmitJobs = 1000
    sacctmgr: sacctmgr:  Modified user associations...
      C = micro      A = account2             U = user5    
      C = micro      A = account2             U = user4    
      C = micro      A = account1             U = user3    
      C = micro      A = account1             U = user2    
      C = micro      A = account1             U = user1    
      C = micro      A = account0             U = admin    
      C = micro      A = root                 U = root     
    sacctmgr: sacctmgr: sacctmgr: Account                 Cluster       User     Share 
    -------------------- ---------- ---------- --------- 
    root                      micro                    1 
     root                     micro       root         1 
     account0                 micro                  100 
      account0                micro      admin         1 
     account1                 micro                  100 
      account1                micro      user1         1 
      account1                micro      user2         1 
      account1                micro      user3         1 
     account2                 micro                  100 
      account2                micro      user4         1 
      account2                micro      user5         1 
    sacctmgr: 
    [INFO] Launching slurmctld
    ['/opt/slurm_sim/sbin/slurmctld', '-e', '/home/slurm/work/micro_cluster/job_traces/small.events', '-dtstart', '59']
    [INFO] Current time 1710513133.6579173
    [INFO] slurmdbd_create_time=1710513112.42
    [INFO] slurmctld_create_time=1710513128.36
    [INFO] slurmd_create_time=1710513127.36
    [INFO] Starting job submittion
    [INFO] Monitoring slurmctld until completion
    [INFO] All jobs submitted wrapping up
    [INFO] slurmctld took 30.056071043014526 seconds to run.
    first_line [2022-01-01T05:00:15.938834] error: Unable to open pidfile `/var/run/slurmctld.pid': Permission denied
     2022-01-01 05:00:15.938834 2022-01-01 05:00:15.938834
    last_line [2022-01-01T05:03:57.912788] All done.
     2022-01-01 05:03:57.912788 2022-01-01 05:03:57.912788
    [INFO] Copying results to :/home/slurm/work/micro_cluster/results/slurmsimcont/test1/dtstart_59_1
    [INFO] copying resulting file /home/slurm/work/micro_cluster/log/jobcomp.log to /home/slurm/work/micro_cluster/results/slurmsimcont/test1/dtstart_59_1
    [INFO] copying resulting file /home/slurm/work/micro_cluster/log/slurmctld.log to /home/slurm/work/micro_cluster/results/slurmsimcont/test1/dtstart_59_1
    [INFO] copying resulting file /home/slurm/work/micro_cluster/log/sched.log to /home/slurm/work/micro_cluster/results/slurmsimcont/test1/dtstart_59_1
    [DEBUG] Submit time for first job: 2022-01-01 05:01:14.315371
    [INFO] copying resulting file /home/slurm/work/micro_cluster/log/slurmctld_stdout.log to /home/slurm/work/micro_cluster/results/slurmsimcont/test1/dtstart_59_1
    [INFO] copying resulting file /home/slurm/work/micro_cluster/log/slurmdbd_stdout.log to /home/slurm/work/micro_cluster/results/slurmsimcont/test1/dtstart_59_1
    [INFO] Simulated time: 0:03:41.973954
    [INFO] Real time: 0:00:30.056383
    [INFO] Acceleration: 7.385252
    [INFO] Done
```

While it is running, we can take a look at specified arguments.
First, we started with setting some bash variables, which will be convenient later when we automate runs for multiple simulations.
We will go through all of them and describe what it is and why we are doing it in this way.

```bash
export CLUS_DIR=$(pwd)
```

It is convenient to specify pathways relative to our cluster simulation directory:

```bash
export MACHINE_NAME="slurmsimcont"
```
IT specidy machine name. 
Just like with real Slurm, the speed of actual hardware can affect the scheduling, so it is helpful to track which hardware was used for simulation.

```bash
export RUN_NAME="test1"
```

`RUN_NAME` is used to label a particular test, for example, control, high-priority, and so on.

```bash
export dtstart=59
export replica=1
```

Due to *stochasticity*, we need to have multiple runs. 
We track them with `replica` parameter for identical run and `dtstart` for additional randomization.

`replica` parameter is usually having values like 1,2,3,....

`dtstart` specifies the time delay between the slurm controller start-up and the first job submission.
It is the main randomization mechanism in Slurm Simulator.


The following arguments were used for `slurmsim` CLI:
* `-v` for extra messages
* `run_sim` command to run single Slurm simulation.
* `-d` remove results from the previous simulation
* `-e` specify the location of slurm `etc.` directory `slurm.conf` and friends should be where
* `-a` specify the sacctmgr script to set up accounts, users, qos and other things
* `-w` specify workload event file (a.k.a. job traces file)
*  `-r` specify results storage directory, we use following format:
  `${CLUS_DIR}/results/${MACHINE_NAME}/${RUN_NAME}/dtstart_${dtstart}_${replica}`
* `-dtstart` specifies the time delay between the slurm controller start-up and the first job submission.


The `run_sim` command simplifies the execution of the Slurm Simulator. It does the following:

1.  With `-d` option, it removes results from the previous simulation (clean DB, remove logs, shared memory, etc)
2. Check the configuration files; some of the options are not supported in the Simulator, and some others should be set to a particular value. The configuration checking helps to run the simulation. Note that not everything is checked.
3. Create missing directories
4. Launch `slurmdbd`
5. Execute `sacctmgr` script
6. Launch `slurmctld` and start simulation
7. On finishing `slurmctld` process termenate `slurmdbd`.
8. Prepocess some outputs so that they can be loaded to R.
9. Copy all resulting files to the results directory (set with `-r` option)


Now take a closer look at the output, you should be able to see the above steps. At the end you will see something like this:
```
[INFO] Simulated time: 0:03:41.973954
[INFO] Real time: 0:00:30.056383
[INFO] Acceleration: 7.3852
```

This informs you on simulated time and the time acceleration.

Take look at result directory:

```{bash}
ls results/slurmsimcont/test1/dtstart_59_1
```

```         
jobcomp.log
perf_profile.log
perf_stat.log
sched.log
slurm_acct.out
slurmctld.log
slurmctld_log.csv
slurmctld_stdout.log
slurmdbd_stdout.log
```

There are a lot of logs and output files:

-   `sched.log` - scheduler log, usually emptyl
-   `slurm_acct.out` - `sacct` output at the end of the simulation
-   `jobcomp.log` - log from `jobcomp` plug-in
-   `slurmctl.log` - log from `slurmctld`
-   `slurmctld_stdout.log` - standard output and standard error from
    `slurmctld`, useful if simulation failed
-   `slurmdbd_stdout.log` - standard output and standard error from `slurmdbd`
    daemon, useful if simulation failed
-   `slurmctld_log.csv` - Slurm controller events generated from `slurmctld.log`
    processing

Lets take a look at `slurm_acct.out`

``` bash
head -n 5 results/slurmsimcont/test1/dtstart_59_1/slurm_acct.out
```

```         
    JobID|JobIDRaw|Cluster|Partition|Account|Group|GID|User|UID|Submit|Eligible|Start|End|Elapsed|ExitCode|State|NNodes|NCPUS|ReqCPUS|ReqMem|ReqTRES|Timelimit|QOS|NodeList|JobName|NTasks
    1001|1001|micro|normal|account2|slurm|1000|user5|1005|2022-01-01T05:01:14|2022-01-01T05:01:14|2022-01-01T05:01:14|2022-01-01T05:01:14|00:00:00|0:0|COMPLETED|1|12|12|33600M|billing=12,cpu=12,mem=33600M,node=1|00:01:00|normal|b1|jobid_1001|
    1002|1002|micro|normal|account1|slurm|1000|user1|1001|2022-01-01T05:01:15|2022-01-01T05:01:15|2022-01-01T05:01:16|2022-01-01T05:02:16|00:01:00|0:0|TIMEOUT|1|1|1|2800M|billing=1,cpu=1,mem=2800M,node=1|00:01:00|normal|n1|jobid_1002|
    1003|1003|micro|normal|account2|slurm|1000|user4|1004|2022-01-01T05:01:16|2022-01-01T05:01:16|2022-01-01T05:01:16|2022-01-01T05:01:21|00:00:05|0:0|COMPLETED|1|1|1|500000M|billing=1,cpu=1,mem=500000M,node=1|00:01:00|normal|b1|jobid_1003|
    1004|1004|micro|normal|account1|slurm|1000|user3|1003|2022-01-01T05:01:30|2022-01-01T05:01:30|2022-01-01T05:01:30|2022-01-01T05:01:51|00:00:21|0:0|COMPLETED|2|24|24|67200M|billing=24,cpu=24,mem=67200M,node=1|00:01:00|normal|b1,g1|jobid_1004|
```

It contains information on compleded jobs. It will be used for our analysis.

You also can find two new directories `log` and `var`. Thouse are used temporary
while Slurm Simulator is running all importent files are eventually copied to
results directory.

We rarely would have only one simulation (need statistics), so before moving to
results reading lets make another simulation:

``` bash
export CLUS_DIR=$(pwd)
export MACHINE_NAME="slurmsimcont"
export RUN_NAME="test1"
export dtstart=79
export replica=1

slurmsim -v run_sim  -d \
            -e ${CLUS_DIR}/etc \
            -a ${CLUS_DIR}/etc/sacctmgr.script \
            -w ${CLUS_DIR}/workload/small.events \
            -r ${CLUS_DIR}/results/${MACHINE_NAME}/${RUN_NAME}/dtstart_${dtstart}_${replica} \
            -dtstart $dtstart
```

## Read Results

Because there is a need to handle multiple runs at a same time we have developed a 
tools which help us with that.
`read_sacct_out_multiple` will read multiple `slurm_acct.out` from simulations 
with different start time and replicas.

```{r}
sacct <- read_sacct_out_multiple(
    slurm_mode="test1", # name of simulation
    results_root_dir="results/slurmsimcont/test1",
    dtstart_list=c(59, 79), # start time list
    run_id_list=1, # replicas list
    # sacct_out="slurm_acct.out"  # non-standard name of sacct_out
)

sacct
```
`read_sacct_out_multiple` also recalculates `SubmitTime`, `StartTime` and `EndTime` 
in reference to the submission of first job.

`read_events_multiple` read events which ever extracted from slurmctrl logs:

```{r}
events_time <- read_events_multiple(
    slurm_mode="test1", # name of simulation
    results_root_dir="results/slurmsimcont/test1",
    dtstart_list=c(59, 79), # start time list
    run_id_list=1, # replicas list
    #events_csv="slurmctld_log.csv" # non-standard name of slurmctld_log.csv
)
events_time
```
## Make Events Plot

Let's make a plot for starting times of our two simulations

```{r}
plot_grid(
  ggplot(sacct, aes(
    x=SubmitTime,y=JobRecID,color=factor(dtstart),shape=factor(dtstart)))+
    geom_point() + scale_shape_manual(values = c(3,4)),
  ggplot(sacct, aes(
    x=StartTime,y=JobRecID,color=factor(dtstart),shape=factor(dtstart)))+
    geom_point() + scale_shape_manual(values = c(3,4)),
  labels = c("A","B"), nrow=2
)

```
You can find that even though submit time is same between two realization the start time can be substantially different.

What are the reasons for such behavior? 
Many Slurm routines are executed in cyclic manner: some will go to sleep for 
predefined amount of time before repeating the cycle, others will check time to time
was a predefined amount of time passed since the last time cycle was started.

For example the function that 
kills jobs running over the requested walltime, start a new cycle 
if 30 seconds passed from last run and then it willcheck all jobs. 
The thread which do the job also do other things so time between checks is not always exact 30 seconds.

In addition we don't know a-priori. at which stage of these varying stop and start cycles the job submission ended up.
So we have to try all different possibilities and report an average behaiviour.

To identify what exactly went different we can use event diagramm:

```{r}


```




